<<<<<<< HEAD
---
title: "Investigating COVID-19 Virus Trends"
output: github_document
---

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

In the solidarity to face this unprecedented global crisis, several organizations did not hesitate to share several datasets allowing the conduction of several kinds of analysis in order to understand this pandemic.

Here, I use [a dataset, from Kaggle](https://www.kaggle.com/lin0li/covid19testing), that **Dataquest** has prepared and made [available here for download](https://dq-content.s3.amazonaws.com/505/covid19.csv). This dataset was collected between the 20th of January and the 1st of June 2020.

This analysis tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

## Understanding the Data
### Loading the Data

```{r}
library(readr)

# Loading the dataset
covid_df = read.csv("sources/covid19.csv")
```

After loading the dataset, we can explore it:

```{r}
# Displaing the dimension of the data: 
dim(covid_df)

# Storing the column names in a variable
vector_cols <- colnames(covid_df)

# Displaing the variable vector_cols
vector_cols

# Showing the first few rows of the dataset
head(covid_df)

# Showing a global view of the dataset.
library(tibble)
glimpse(covid_df)
```

Our dataset contains the following columns:\
1.`Date`: Date\
2. `Continent_Name`: Continent names\
3. `Two_Letter_Country_Code`: Country codes\
4. `Country_Region`: Country names\
5. `Province_State`: States/province names; value is `All States` when state/provincial level data is not available\
6. `positive`: Cumulative number of positive cases reported.\
7. `active`: Number of active cases on that **day**.\
8. `hospitalized`: Cumulative number of hospitalized cases reported.\
9. `hospitalizedCurr`: Number of actively hospitalized cases on that **day**.\
10. `recovered`: Cumulative number of recovered cases reported.\
11. `death`: Cumulative number of deaths reported.\
12. `total_tested`: Cumulative number of tests conducted.\
13. `daily_tested`: Number of tests conducted on the **day**; if daily data is unavailable, daily tested is averaged across number of days in between.\
14. `daily_positive`: Number of positive cases reported on the **day**; if daily data is unavailable, daily positive is averaged across number of days in.\


## Isolating the Rows We Need

Looking at the few lines of our dataset we displayed in the previous step, we can see that the `Province_State` column mixes data from different levels: country level and state/province level. Since we cannot run an analysis on all these levels at the same time, we need to filter what we are interested in.

We will, therefore, extract only the country-level data in order to not bias our analyses. To do so, we filter the data to keep only the data related to `"All States"`. `"All States"` represents the value of the column Province_State to specify that the COVID-19 data is only available at the country level.

```{r}
library(dplyr)
covid_df_all_states <- covid_df %>% filter(Province_State == "All States") %>% select(-Province_State)

# Data glimpse
head(covid_df_all_states)
```

## Isolating the Columns We Need

Revisiting the description of the dataset columns, we can notice that there are columns that provide daily information and others that provide cumulative information. 

Hence, we should manage those cases (columns with cumulative and daily information) separately because we cannot work with both together. Actually, our analysis would be biased if we made the mistake of comparing a column containing cumulative data and another one containing only one-day data.

Thereafter, we work mainly with daily data. So let's extract the columns related to the daily measures.

```{r}
covid_df_all_states_daily <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

# Data glimpse
head(covid_df_all_states_daily)
```

## Extracting the Top Ten Tested Cases Countries

Our goal here is to extract the top ten cases countries data. Acting like a data scientist, at this step, these are the questions we are asking ourselves.

- How can we get the overall number of COVID-19 tested, positive, active and hospitalized cases by country since we currently have daily data?

- How do we then extract the top ten?

First of all, we group the dataset by the `Counrtry_Region` column and compute the sum of the number of tested, positive, active and hospitalized cases and arrange the dataframe accordingly:

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(tested=sum(daily_tested), positive=sum(daily_positive), active=sum(active), hospitalized=sum(hospitalizedCurr)) %>% arrange(-tested)

# Visualisation of generated dataframe:
covid_df_all_states_daily_sum
```

Now, we can extract the top-ten countries:

```{r}
covid_top_10 <- covid_df_all_states_daily_sum %>% head(10)
```


## Identifying the Highest Positive Against Tested Cases

As a remainder, our goal is to answer this question: ** Which countries have had the highest number of positive cases against the number of tests?**

Therefore we extract the columns as vectors:

```{r}
# extract vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases  <- covid_top_10$positive
active_cases  <- covid_top_10$active
hospitalized_cases  <- covid_top_10$hospitalized

# name the values of the vectors
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

Now, we can identify the top three positive against tested cases:

```{r}
positive_against_tested <- positive_cases / tested_cases

positive_against_tested_sorted <- sort(positive_against_tested, decreasing= TRUE)

positive_tested_top_3 <- positive_against_tested_sorted[1:3]

positive_tested_top_3
```

## Keeping relevant information

Our goal is to find a way to keep all the information available for the top three countries that have had the highest number of positive cases against the number of tests.

The previous step allowed identifying those top three countries as:

```{r}
positive_tested_top_3
```

To make sure we won't lose other information about these countries we can create a matrix that contains the ratio and the overall number of COVID-19 tested, positive, active and hospitalized cases. 

```{r}
# creating country vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

# creating matrix from vectors
covid_mat <- rbind(united_kingdom,united_states,turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat
```

## Putting all together

Here, our  goal is to put all our answers and datasets together. Since a list can contain several types of objects, we are able to store all the data of our project together. This allows us to have a global view from a single variable and the ability to export our results for other uses.

```{r}
# Create a character variable that contains our question
question <- "Which countries have had the highest number of positive cases against the number of tests?"

# Create a character variable that contains our answer
answer <- c("Positive tested cases" = positive_tested_top_3)

# Create lists that contain the data structures we created
dataframe_list <- list(original = covid_df,
                       allstates = covid_df_all_states,
                       daily = covid_df_all_states_daily,
                       top_10 = covid_top_10)
matrix_list <- list(covid_mat)
vector_list <- list(vector_cols,countries)

# Create a named list that contains the three previous lists associated with the data structure names
data_structure_list <- list("dataframe" = dataframe_list,
                            "matrix" = matrix_list,
                            "vector" = vector_list)

# Create a list that summarizes our results
covid_analysis_list <- list(question,answer,data_structure_list)

# Display the second element
covid_analysis_list[[2]]
```

=======
---
title: "Investigating COVID-19 Virus Trends"
output: github_document
---

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

In the solidarity to face this unprecedented global crisis, several organizations did not hesitate to share several datasets allowing the conduction of several kinds of analysis in order to understand this pandemic.

Here, I use [a dataset, from Kaggle](https://www.kaggle.com/lin0li/covid19testing), that **Dataquest** has prepared and made [available here for download](https://dq-content.s3.amazonaws.com/505/covid19.csv). This dataset was collected between the 20th of January and the 1st of June 2020.

This analysis tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

## Understanding the Data
### Loading the Data

```{r}
library(readr)

# Loading the dataset
covid_df = read.csv("sources/covid19.csv")
```

After loading the dataset, we can explore it:

```{r}
# Displaing the dimension of the data: 
dim(covid_df)

# Storing the column names in a variable
vector_cols <- colnames(covid_df)

# Displaing the variable vector_cols
vector_cols

# Showing the first few rows of the dataset
head(covid_df)

# Showing a global view of the dataset.
library(tibble)
glimpse(covid_df)
```

Our dataset contains the following columns:\
1.`Date`: Date\
2. `Continent_Name`: Continent names\
3. `Two_Letter_Country_Code`: Country codes\
4. `Country_Region`: Country names\
5. `Province_State`: States/province names; value is `All States` when state/provincial level data is not available\
6. `positive`: Cumulative number of positive cases reported.\
7. `active`: Number of active cases on that **day**.\
8. `hospitalized`: Cumulative number of hospitalized cases reported.\
9. `hospitalizedCurr`: Number of actively hospitalized cases on that **day**.\
10. `recovered`: Cumulative number of recovered cases reported.\
11. `death`: Cumulative number of deaths reported.\
12. `total_tested`: Cumulative number of tests conducted.\
13. `daily_tested`: Number of tests conducted on the **day**; if daily data is unavailable, daily tested is averaged across number of days in between.\
14. `daily_positive`: Number of positive cases reported on the **day**; if daily data is unavailable, daily positive is averaged across number of days in.\


## Isolating the Rows We Need

Looking at the few lines of our dataset we displayed in the previous step, we can see that the `Province_State` column mixes data from different levels: country level and state/province level. Since we cannot run an analysis on all these levels at the same time, we need to filter what we are interested in.

We will, therefore, extract only the country-level data in order to not bias our analyses. To do so, we filter the data to keep only the data related to `"All States"`. `"All States"` represents the value of the column Province_State to specify that the COVID-19 data is only available at the country level.

```{r}
library(dplyr)
covid_df_all_states <- covid_df %>% filter(Province_State == "All States") %>% select(-Province_State)

# Data glimpse
head(covid_df_all_states)
```

## Isolating the Columns We Need

Revisiting the description of the dataset columns, we can notice that there are columns that provide daily information and others that provide cumulative information. 

Hence, we should manage those cases (columns with cumulative and daily information) separately because we cannot work with both together. Actually, our analysis would be biased if we made the mistake of comparing a column containing cumulative data and another one containing only one-day data.

Thereafter, we work mainly with daily data. So let's extract the columns related to the daily measures.

```{r}
covid_df_all_states_daily <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

# Data glimpse
head(covid_df_all_states_daily)
```

## Extracting the Top Ten Tested Cases Countries

Our goal here is to extract the top ten cases countries data. Acting like a data scientist, at this step, these are the questions we are asking ourselves.

- How can we get the overall number of COVID-19 tested, positive, active and hospitalized cases by country since we currently have daily data?

- How do we then extract the top ten?

First of all, we group the dataset by the `Counrtry_Region` column and compute the sum of the number of tested, positive, active and hospitalized cases and arrange the dataframe accordingly:

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(tested=sum(daily_tested), positive=sum(daily_positive), active=sum(active), hospitalized=sum(hospitalizedCurr)) %>% arrange(-tested)

# Visualisation of generated dataframe:
covid_df_all_states_daily_sum
```

Now, we can extract the top-ten countries:

```{r}
covid_top_10 <- covid_df_all_states_daily_sum %>% head(10)
```


## Identifying the Highest Positive Against Tested Cases

As a remainder, our goal is to answer this question: ** Which countries have had the highest number of positive cases against the number of tests?**

Therefore we extract the columns as vectors:

```{r}
# extract vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases  <- covid_top_10$positive
active_cases  <- covid_top_10$active
hospitalized_cases  <- covid_top_10$hospitalized

# name the values of the vectors
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

Now, we can identify the top three positive against tested cases:

```{r}
positive_against_tested <- positive_cases / tested_cases

positive_against_tested_sorted <- sort(positive_against_tested, decreasing= TRUE)

positive_tested_top_3 <- positive_against_tested_sorted[1:3]

positive_tested_top_3
```

## Keeping relevant information

Our goal is to find a way to keep all the information available for the top three countries that have had the highest number of positive cases against the number of tests.

The previous step allowed identifying those top three countries as:

```{r}
positive_tested_top_3
```

To make sure we won't lose other information about these countries we can create a matrix that contains the ratio and the overall number of COVID-19 tested, positive, active and hospitalized cases. 

```{r}
# creating country vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

# creating matrix from vectors
covid_mat <- rbind(united_kingdom,united_states,turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat
```

## Putting all together

Here, our  goal is to put all our answers and datasets together. Since a list can contain several types of objects, we are able to store all the data of our project together. This allows us to have a global view from a single variable and the ability to export our results for other uses.

```{r}
# Create a character variable that contains our question
question <- "Which countries have had the highest number of positive cases against the number of tests?"

# Create a character variable that contains our answer
answer <- c("Positive tested cases" = positive_tested_top_3)

# Create lists that contain the data structures we created
dataframe_list <- list(original = covid_df,
                       allstates = covid_df_all_states,
                       daily = covid_df_all_states_daily,
                       top_10 = covid_top_10)
matrix_list <- list(covid_mat)
vector_list <- list(vector_cols,countries)

# Create a named list that contains the three previous lists associated with the data structure names
data_structure_list <- list("dataframe" = dataframe_list,
                            "matrix" = matrix_list,
                            "vector" = vector_list)

# Create a list that summarizes our results
covid_analysis_list <- list(question,answer,data_structure_list)

# Display the second element
covid_analysis_list[[2]]
```

>>>>>>> 87053f14808bb3121a766074450f4b2cd9f8096d
