---
title: "Creating An Efficient Data Analysis Workflow"
author: "Bastian Hartmann"
date: "22 Januar 2022"
output: html_document
---

In this project, we will be acting as a data analyst for a company that sells books for learning programming. Our company has produced multiple books, and each has received many reviews. Our company wants us to check out the sales data and see if we can extract any useful information from it.

The dataset of textbook sales we use, can be downloaded [here](https://data.world/dataquest/book-reviews).

To start, we first load the dataset and get familiar with it.

```{r, include=FALSE}
# This code section is for loading libraries:
library(readr)
library(dplyr)
```


## Getting Familiar With The Data

It's easy to lose context when we're just talking about data analysis in general. The first thing we should do before we do any analysis is to get acquainted with our dataset. There are many, many things to check with a dataset before we dive into any analysis. How much data is there? What kind of data do we actually have on hand? Is there anything "weird" that might interfere with any analyses we might need to do? Is there missing data? Answering these questions now saves we time and effort later.

If we don't check the data beforehand, it's easy to make some false assumptions about the data that can hinder our progress later.

### Loading the Dataset

```{r}
reviews <- read_csv("resources/book_reviews.csv")
col_names <- colnames(reviews)
head(reviews)
```

```{r}
dim(reviews)
```

Here, we can see that the dataset consists of 2000 rows and 4 columns.

The four columns are:

1. `book` -> Title of the book. [Type: character]
2. `review` -> The review of the sold book in word terms like "Excellent", "Fair", "Poor", etc.. [Type: character]
3. `state` -> The state where the book has been sold. [Type: character]
4. `price` -> Price of the book as numeric values in US-Dollar. [Type: double]

Next, we want to observe the unique values and their counts of each columns.

### The `book` Column

```{r}
unique_book <- count(reviews,book,sort=TRUE)
unique_book
```

In this summary of the `book` column, we can see that in the dataset are five different books contained. Each has about 400 occurences.

### The `review` column

```{r}
unique_review <- count(reviews,review,sort=TRUE)
unique_review
```

Here, we can see that there are five different unique reviews. In the dataset are also 206 entries without an review (NA values).

### The `state` column

```{r}
unique_state <- count(reviews,state,sort=TRUE)
unique_state
```

We have 8 different states in the dataset mensioned in the `state` column.

### The `price` column

```{r}
unique_price <- count(reviews,price,sort=TRUE)
unique_price
```
Each of our five book has an unique price value.

The prices range from US\$15.99 to US\$50.00.

## Handling Missing Data

We will take care of the missing values in the `review` column by removing the rows with a `NA` value.

```{r}
filtered_reviews <- reviews %>% filter((!is.na(reviews[["review"]])))
dim(filtered_reviews)
```

From the dimensions of `filtered_reviews`, we can see that the 206 rows with the `NA` value are now dropped. (2000 - 206 = 1794)

## Transforming The Review Data

Our goal is to evaluate the ratings of each of the textbooks, but there's not much we can do with text versions of the review scores. It would be better if we were to convert the reviews into a numerical form. Thus, we create the new column `review_num` containing number from `1` to `5` corresponding to the reviews from `"Poor"` to `"Excellent"`.

```{r}
filtered_reviews_numReviews <- filtered_reviews %>% mutate(
  review_num = case_when(
    review == "Poor" ~ 1,
    review == "Fair" ~ 2,
    review == "Good" ~ 3,
    review == "Great" ~ 4,
    review == "Excellent" ~ 5
  )
)

head(filtered_reviews_numReviews)
```

Next, we also whant to have a column (`is_high_review`) that helps us decide if a score is "high" (4 or higher) or not.

```{r}
filtered_reviews_numAndHigh <-filtered_reviews_numReviews %>% mutate(
  is_high_review = if_else(review_num >= 4,TRUE,FALSE)
)

head(filtered_reviews_numAndHigh)
```

## Analyzing The Data

It's important to keep the overall goal in mind as we handle all the little details of the cleaning. We are acting as an analyst trying to figure out which books are the most profitable for the company.

Our main goal is to figure out what book is the most profitable. We will judge this by calculating how much money each book generates overall.

```{r}
book_revenues <- filtered_reviews_numAndHigh %>% group_by(book) %>% summarize(Revenue = sum(price), N = n()) %>% arrange(desc(Revenue))

book_revenues
```

## Conclusion

Using our score, the most profitable book is **"Secrets Of R For Advanced Students"**.