---
title: "NYC Schools Perceptions"
author: "Bastian Hartmann"
date: "05 February 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

In this project, we'll investigate the responses to [surveys](https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8) designed to gauge parent, student, and teacher perceptions of the quality of New York City schools. 

More precisely, we want to investigate the following questions:

- Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?
- Do students, teachers, and parents have similar perceptions of NYC school quality?

The data we use, can be downloaded [here](https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8). The `.zip` folder contains five files:

- `Survey Data Dictionary.xls`
- `masterfile11_gened_final.xlsx`
- `masterfile11_gened_final.txt`
- `masterfile11_d75_final.xlsx`
- `masterfile11_d75_final.txt`

`Survey Data Dictionary.xls`  contains metadata that will be useful as you decide how to clean and prepare the survey data for analysis.

`masterfile11_gened_final.xlsx` and `masterfile11_gened_final.txt` contain survey data for "general education" schools — those that do not specifically serve populations with special needs.

`masterfile11_d75_final.xlsx` and `masterfile11_d75_final.txt` contain survey data for [District 75](https://www.schools.nyc.gov/special-education/school-settings/district-75) schools, which provide exceptional education support for children with special needs such as learning or physical disabilities.

In addition, we also use the `"combined.csv"` file that can be downloaded [here](https://data.world/dataquest/nyc-schools-data/workspace/file?filename=combined.csv). This file contains various information about NYC schools like SAT scores and ethnicity.

## Loading the Survey Data

First, we have to import the data:

```{r}
combined <- read_csv("sources/combined.csv",show_col_types = FALSE)
survey_gened <- read_tsv("sources/2011 data files online/masterfile11_gened_final.txt",show_col_types = FALSE)
survey_d75 <- read_tsv("sources/2011 data files online/masterfile11_d75_final.txt",show_col_types = FALSE)
```

To get an overview of the dataset, let's have a look at the first 10 rows of each:

```{r}
knitr::kable(head(combined))
knitr::kable(head(survey_gened))
knitr::kable(head(survey_d75))
```

In addition, the `Survey Data Dictionary.xls` files contains the following information about the "Field Names", "Field Series" and "Field Convention" survey data:

```{r}
library(readxl)
Field_Names <- read_excel("sources/2011 data files online/Survey Data Dictionary.xls", 
    range = "A4:B36")
Field_Series <- read_excel("sources/2011 data files online/Survey Data Dictionary.xls", 
    range = "A37:B46")
Field_Convention <- read_excel("sources/2011 data files online/Survey Data Dictionary.xls", 
    range = "A47:B59")
```

```{r}
knitr::kable(Field_Names)
knitr::kable(Field_Series)
knitr::kable(Field_Convention)
```

---

## Simplifying the Dataframes

After observing variables in the survey data, we can see that the variable names refer to schools' scores as determined by different groups' survey responses.

Survey questions assess the following metrics:

- Safety and Respect
- Communication
- Engagement
- Academic Expectations

The following groups responded to the survey:

- Parents
- Teachers
- Students
- Total (The average of parent, teacher, and student scores)

We can significantly simplify the dataframes and make them easier to work with by creating new dataframes that contain only variables aggregating scores based on survey results.

First, we only want to include variables we will need for our analysis. Thus, for the `survey_gened` dataset we will drop all rows *not* containing data from "High School" and include only the columns from `dbn` to `aca_tot_11` (varaibles that contain aggregate scores). The same column filtering will be applied to the `survey_d75` dataset.

```{r}
survey_select <- survey_gened %>%
  filter(schooltype=="High School") %>%
  select(dbn:aca_tot_11)

survey_d75_select <- survey_d75 %>%
  select(dbn:aca_tot_11)
```

---

## Creating a Single Dataframe for Analyis

Now that you've simplified them, working with the survey dataframes probably feels more manageable. Next, we want to join the survey data to the `combined` dataframe containing NYC school demographic and test score data.

Therefore, we first join the two survey dataframes:

```{r}
survey_data <- survey_select %>%
  bind_rows(survey_d75_select)
```

Next, we will join the `survey_data` and the `combined` dataframes. Therefore, we rename the `dbn` column to `DBN` of the `survey_data`, so it can be used as key to join with the `combined` data frame.

```{r}
survey_data <- survey_data %>%
  rename(DBN=dbn)
```

Now, we join the `combined` and `survey_total` dataframes using a *left join* to keep only survey data that correspond to schools for which we have data in `combined`.

```{r}
complete_survey <- combined %>%
  left_join(survey_data,by="DBN")
```

---

## Investigating Interesting Correlations and Relationships

Now that you've created a single, clean dataframe to work with, we can begin with our analysis.

Our first question we want to investigate was:

- **Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?**


To get an idea of which demographic and test score variables may be related to parent, teacher, and student perceptions of NYC school quality, we'll make a **correlation matrix** of the numeric variables with the `avg_sat_score`. Furthermore, we filter for correlation coefficients higher than 0.25 or lower than -0.25 to identify possible strong correlations and sort them in descending order of their absolute value.

```{r Correlation_matrix_of_demographic_and_score_variables, warning=FALSE}
cor_mat <- complete_survey %>%  
  select(where(is.numeric)) %>%
  cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%#
  as_tibble(rownames="variable")

SATscore_cors <- cor_tib %>%
  select(variable,avg_sat_score) %>%
  filter(avg_sat_score > 0.25 | avg_sat_score < -0.25) %>%
  mutate(abs_score_cors=abs(avg_sat_score)) %>%
  arrange(desc(abs_score_cors)) %>%
  select(-abs_score_cors) %>%
  mutate(variable=str_replace_all(variable," ","_"))

knitr::kable(SATscore_cors)
```

The top five strong correlations, that are not directly connected to the SAT score are:

- frl_percent (-0.7235302): 
- white_per (0.6495892): 
- Total_Exmas_Taken (0.6071524):
- asian_per (0.5659381): 
- dropout_precent (-0.4821499): 

To further examine the correlation, we make scatter plots of the these five variables vs. `avg_sat_score`:

```{r Scatter_plots, warning=FALSE}
display_scatter <- function(x,y) {
  ggplot(data=complete_survey ,aes_string(x=x,y=y))+
  geom_point()+
  ggtitle(x)
}

x_vars <- SATscore_cors$variable[c(5,6,8,9,13)]
y_var <- "avg_sat_score"

column_names <- colnames(complete_survey)
column_names <- str_replace_all(column_names," ","_")
colnames(complete_survey) <- column_names

map2(x_vars,y_var,display_scatter)
```

The scatter plots confirm the strong correlations of the selected variables with the `avg_sat_score` variable. Notably, this shows that the average SAT score is better when more white students (`white_per`), more asian students (`asian_per`) are at a High School and how many total exams are taken (`Total_Exams_Taken`). In contrast, the average SAT score shows a strong negative correlation with the number of offered free and reduced lunches (`frl_percent`) and the number of dropouts (`dropout_percent`).

---

## Differences in Student, Parent, and Teacher Perceptions

We've worked on exploring potential relationships between perspectives of NYC school quality and demographic and academic success metrics.

Furthermore, it  would be interesting to see whether parents, students, and teachers have similar perceptions about the four school quality metrics they were surveyed about:

- Safety and Respect
- Communication
- Engagement
- Academic Expectations

To address this question, we first have to reshape the dataframe to have better visualization possibilities. We start by reshaping the dataframe that the survey questions (like `saf_p_11` and `com_s_11`) and the score (like `8.7` or `9.0`) are in two separate columns.

```{r}
survey_data_long <- complete_survey %>%
  pivot_longer(
    cols = colnames(complete_survey[46:61]),
    names_to = "Field_Names",
    values_to = "Scores"
  ) %>%
  drop_na(Scores)
```

After this step, we will create two new variables (`response_type` and `metric`).

- `response_type`: This variable indicates the type of the survey participant - "parent", "teacher", "student", or "total".
- `metric`: This variable indicates the survey category - "engagement", "academic expectations", "safety and respect", or "communication".

```{r}
det_response_type <- function(field_name){
  if (str_detect(field_name,"_p_")){
    return ("parent")
  } else if (str_detect(field_name,"_t_")){
    return ("teacher")
  } else if (str_detect(field_name,"_s_")){
    return ("student")
  } else if (str_detect(field_name,"_tot_")){
    return ("total")
  }
}

det_metric_type <- function(field_name){
  if (str_starts(field_name,"eng")){
    return ("engagement")
  } else if (str_starts(field_name,"aca")){
    return ("academic expectations")
  } else if (str_starts(field_name,"saf")){
    return ("safety and respect")
  } else if (str_starts(field_name,"com")){
    return ("communication")
  }
}

survey_data_long <- survey_data_long %>%
  mutate(
    response_type = map(Field_Names,det_response_type),
    metric = map(Field_Names,det_metric_type)
  )

survey_data_long$response_type <- factor(survey_data_long$response_type,levels = c("parent","teacher","student","total"))

survey_data_long$metric <- factor(survey_data_long$metric,levels = c("engagement","academic expectations","safety and respect","communication"))
```

Now, we are ready to visualize differences in parent, teacher, and student perceptions of NYC school quality metrics using a box plot.

```{r Box_Plot_Response_Scores, fig.height=8, fig.width=10}

survey_data_long %>%
  ggplot(aes(x=response_type,y=Scores,fill=response_type)) +
  geom_boxplot()+
  labs(
    title = "Response Scores of the Survey Groups to Different Metrics",
    x = "Score Metric",
    fill = "Respons Group" 
  )+
  facet_wrap(facets = vars(metric),
             nrow = 2,
             ncol = 2)
```

From these box plots, it looks like the **students** replied in general with the *lowest scores* in all four metrics, but this group also has the highest amount of high outlines. I n contrast, the **parents** are the group that gives the *highest scores* in all four metrics. Whereas, the **teachers** are in the *middle field* regarding their scores but with the biggest spreading (highest IQD).